---
title: "Group 2 - COVID Mobility NYC"
author: "Nirvi Shah and Daniel Glick"
output:
  html_document:
    df_print: paged
---

# Differential COVID-19 case positivity in New York City neighborhoods: Socioeconomic factors and mobility

Replication of figures and tables from the paper https://onlinelibrary.wiley.com/doi/full/10.1111/irv.12816

# Load the Libraries
 
```{r}
library(tidycensus)
library(tidyverse)
library(dplyr)
library(tigris)
library(modelr)
library(sf)

options(tigris_use_cache = TRUE)
Sys.setenv(CENSUS_KEY='1bbcb0a1bcda959d1e5bb3e6b0a04aff304c477f')
# Reload .Renviron
readRenviron("~/.Renviron")
# Check to see that the expected key is output in your R console
Sys.getenv("CENSUS_KEY")
```

# Load the Data

```{r}
zcta <- read_csv("zcta.csv") 
zcta <- zcta %>% transform(MODZCTA = as.factor(MODZCTA))%>% rename(GEOID=MODZCTA) %>% select(GEOID)
varNames <- load_variables(2016, "acs5", cache = TRUE) 
```

# Proportion of the 18- to 64-year-old population that is uninsured - B27010

```{r}
ny_insure <- get_acs(
  geography = "zcta", 
  variables = c('B27010_033','B27010_050','B27010_018','B27010_034'),
  year = 2016,
  geometry = TRUE,
  output = "wide"
)
```

```{r}
ny_insure <- inner_join(ny_insure, zcta, by = "GEOID")
ny_insure <- ny_insure %>% mutate(proportion=(B27010_033E+B27010_050E)/(B27010_018E+B27010_034E))
median(ny_insure$proportion, na.rm = TRUE)
```

```{r}
ggplot(ny_insure, aes(fill = proportion)) +
  ggtitle("Proportion of 18-64 year olds who are uninsured") +
  geom_sf(aes(geometry = geometry)) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  theme(legend.position="bottom")
```


# Median household income - B19013

```{r}
median_income <- get_acs(
  geography = "zip code tabulation area", 
  variables = "B19013_001",
  year = 2016,
  geometry = TRUE,
  output = "wide"
)
```


```{r}
median_income <- inner_join(median_income, zcta, by = "GEOID")
median_income <- median_income %>% mutate(totalincome=(B19013_001E)/1000000)
median(median_income$totalincome, na.rm = TRUE)
```


```{r}
ggplot(median_income, aes(fill = totalincome)) + 
  ggtitle("Median income (in millions, 2016$)") + 
  geom_sf(aes(geometry = geometry)) +
  scale_fill_distiller(palette = "YlGn", direction = 1) + 
  theme(legend.position="bottom")
```

# Proportion of population that self-identified their race as white - B02001

```{r}
white <- get_acs(
  geography = "zip code tabulation area", 
  variables = c("B02001_002","B02001_001"),
  year = 2016,
  geometry = TRUE,
  output = "wide"
)
```


```{r}
white <- inner_join(white, zcta, by = "GEOID")
white <- white %>% mutate(proportion=(B02001_002E)/(B02001_001E))
median(white$proportion, na.rm = TRUE)
```


```{r}
ggplot(white, aes(fill = proportion)) +
  ggtitle("Proportion self-identifying as White") +
  geom_sf(aes(geometry = geometry)) +
  scale_fill_distiller(palette = "Purples", direction = 1) +
  theme(legend.position="bottom")
```

# Proportion of population living in households with more than three inhabitants - B11016

```{r}
household <- get_acs(
  geography = "zip code tabulation area", 
  variables = c("B11016_005","B11016_006","B11016_007","B11016_008","B11016_013","B11016_014","B11016_015","B11016_016","B11016_001"),
  year = 2016,
  geometry = TRUE,
  output = "wide"
)
```


```{r}
household <- inner_join(household, zcta, by = "GEOID")
household <- household %>% mutate(proportion=
          (B11016_005E+B11016_006E+B11016_007E+B11016_008E+
          B11016_013E+B11016_014E+B11016_015E+B11016_016E)
          /(B11016_001E))
median(household$proportion, na.rm = TRUE)
```

```{r}
ggplot(household, aes(fill = proportion)) +
  ggtitle("Proportion self-identifying as White") +
  geom_sf(aes(geometry = geometry)) +
  scale_fill_distiller(palette = "OrRd", direction = 1) +
  theme(legend.position="bottom")
```

# Proportion of population using public transportation to commute to work that includes bus travel - B08301

```{r}
bus <- get_acs(
  geography = "zip code tabulation area", 
  variables = c("B08301_011","B08301_001"),
  year = 2016,
  geometry = TRUE,
  output = "wide"
)
```


```{r}
bus <- inner_join(bus, zcta, by = "GEOID")
bus <- bus %>% mutate(proportion=(B08301_011E)/(B08301_001E))
median(bus$proportion, na.rm = TRUE)
```


```{r}
ggplot(bus, aes(fill = proportion)) +
  ggtitle("Proportion of population that commutes by bus") +
  geom_sf(aes(geometry = geometry)) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  theme(legend.position="bottom")
```

# Proportion of population that is elderly (65+ years of age) - B01001

```{r}
elder <- get_acs(
  geography = "zip code tabulation area", 
  variables = c("B01001_001","B01001_020","B01001_021","B01001_022","B01001_023","B01001_024","B01001_025",
                "B01001_044","B01001_045","B01001_046","B01001_047","B01001_048","B01001_049"),
  year = 2016,
  geometry = TRUE,
  output = "wide"
)
```


```{r}
elder <- inner_join(elder, zcta, by = "GEOID")
elder <- elder %>% mutate(proportion=(B01001_020E+B01001_021E+B01001_022E+B01001_023E+B01001_024E+B01001_025E+
                        B01001_044E+B01001_045E+B01001_046E+B01001_047E+B01001_048E+B01001_049E)/(B01001_001E))
median(elder$proportion, na.rm = TRUE)
```


```{r}
ggplot(elder, aes(fill = proportion)) +
  ggtitle("Proportion of population 65+ years of age") +
  geom_sf(aes(geometry = geometry)) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  theme(legend.position="bottom")
```

```{r}
save(ny_insure, median_income, household, bus, white, elder, file="acs_data.Rdata")
```

# Build Table 1

Build the results shown in Table 1 in the paper

# Load the data

```{r}
load("acs_data.Rdata")

data_april1 <- read_csv("https://raw.githubusercontent.com/nychealth/coronavirus-data/097cbd70aa00eb635b17b177bc4546b2fce21895/tests-by-zcta.csv")

data_april1 <- data_april1 %>% transform(MODZCTA = as.factor(MODZCTA))%>% rename(GEOID=MODZCTA) %>% mutate(positive_proportion = Positive / Total)
```

## Household

```{r}
household_data <- household %>% inner_join(data_april1, by="GEOID")
household_model <- lm(positive_proportion ~ proportion, household_data)
household_data <- add_predictions(household_data, household_model)
cat("In Paper: 41 % \t")
cat("Estimated:", round(rsquare(household_model, household_data) * 100),"%")
```

## Uninsured

```{r}
insure_data <- ny_insure %>% inner_join(data_april1, by="GEOID")
insure_model <- lm(positive_proportion ~ proportion, insure_data)
insure_data <- add_predictions(insure_data, insure_model)
cat("In Paper: 38 % \t")
cat("Estimated:", round(rsquare(insure_model, insure_data) * 100),"%")
```

## White

```{r}
white_data <- white %>% inner_join(data_april1, by="GEOID")
white_model <- lm(positive_proportion ~ proportion, white_data)
white_data <- add_predictions(white_data, white_model)
cat("In Paper: 34 % \t")
cat("Estimated:", round(rsquare(white_model, white_data) * 100),"%")
```

## Median Income

```{r}
income_data <- median_income %>% inner_join(data_april1, by="GEOID")
income_model <- lm(positive_proportion ~ totalincome, income_data)
income_data <- add_predictions(income_data, income_model)
cat("In Paper: 32 % \t")
cat("Estimated:", round(rsquare(income_model, income_data) * 100),"%")
```

## Bus

```{r}
bus_data <- bus %>% inner_join(data_april1, by="GEOID")
bus_model <- lm(positive_proportion ~ proportion, bus_data)
bus_data <- add_predictions(bus_data, bus_model)
cat("In Paper: 13 % \t")
cat("Estimated:", round(rsquare(bus_model, bus_data) * 100),"%")
```

## Elder

```{r}
elder_data <- elder %>% inner_join(data_april1, by="GEOID")
elder_model <- lm(positive_proportion ~ proportion, elder_data)
elder_data <- add_predictions(elder_data, elder_model)
cat("In Paper: 3 % \t")
cat("Estimated:", round(rsquare(elder_model, elder_data) * 100),"%")
```